---
title: "Homework 1 STAT40830-Adv Data Prog with R"
author: "Philip Doonan "
format: pdf
editor: visual
---

[**Student Number:**]{.underline} 24100047

```{r,warning=FALSE,echo=FALSE,message=FALSE}
#load librarys used at start and together
library(ggplot2)
library(dplyr)
library(kableExtra)
```

## [Data set]{.underline}

The data set I have chosen to use is the ***famous iris data*** set which contains different measurements sepal, length and width as well as petal, length and width of three species of the iris flower in ***centimeters***. The three species being ***Iris setosa***, ***versicolor***, and ***virginica***. I have chosen this data set as it contains very distinct clusters between the three species. It is also ***built into R*** and so can be used with any package.

```{r,warning=FALSE,echo=FALSE,message=FALSE}
#load dataset and show first few lines
data(iris)
#table to show data
kable(head(iris),align = "c",
      caption = "First few lines of Iris dataset")|>
  kable_styling(latex_options = c("striped"))
```

## [Plot of data]{.underline}

-   Used ggplot to plot a scatter plot of ***Sepal Length*** and ***Petal length***

-   Set the points colour to the ***species***. To distinguish and show three clusters

-   Added main and axis titles with units(centimeters).

-   Added a theme to improve plots formatting

```{r,warning=FALSE,echo=FALSE,message=FALSE}
#plot data
ggplot(iris)+
  #scatter plot using geom point
  geom_point(aes(x=Sepal.Length,y=Petal.Length,colour=Species))+
  #add labels
  labs(
    title ="Iris flower species petal and sepal length ",
    y='Petal Length (cm)',
    x='Sepal Length (cm)',
  )+
  #add a theme for aesthetic
  theme_light()
```

This figure shows the difference between iris flowers species ***petal*** and ***sepal lengths*** with some overlap in sepal length but very distinguished differences in petal length showing three ***distinct independent clusters.*** Representing the three species as seen with ***setosa*** (Red) in the bottom left with a short petal and sepal lengh, ***versicolor*** with a average petal and sepal length and ***virginica*** with a long petal and sepal length.

```{r,warning=FALSE,echo=FALSE,message=FALSE}
#plot data
ggplot(iris)+
  #scatter plot using geom point
  geom_point(aes(x=Sepal.Width,y=Petal.Width,colour=Species))+
  #add labels
  labs(
    title ="Iris flower species petal and sepal width ",
    y='Petal width (cm)',
    x='Sepal width (cm)',
  )+
  #add a theme for aesthetic
  theme_light()
```
